HO unification from object language to meta language

### 2024-04-01 paper/main.tex
happy with 6, 1 example to double check

\noindent
In the base compilation scheme \elpiIn{progress1} is the identity
on both the links and the substitution, so the fixpoint trivially terminates.
Sections \ref{sec:eta} and \ref{sec:beta} add rules to \elpiIn{progress1}
and justify why the don't hinder termination.

@@

\begin{lemma}
  Forall list of links \elpiIn{L} and \elpiIn{S}, \elpiIn{progress L _ S _}
  terminates
\end{lemma}

\begin{proof}[Proof sketch]
  The addition of rules for \elpiIn{progres1} complicates the function
  \elpiIn{progress}. We can note, however, that they do not prevent the
  termination of \elpiIn{progress}. 1) If a link is activated it is removed from
  \linkStore and the recursive call to \elpiIn{progress} will have a smaller
  list of links to recurse on. Moreover, link activation only runs terminating
  instructions (such as unification). 2) If a link is deduplicated, the
  termination of \elpiIn{progress} is still guaranteed since again we reduce
  \linkStore and the instructions run by link deduplications are all
  terminating. 3) If a link is neither activated nor deduplicated, i.e. it
  remains suspended, then \linkStore remains unchanged like the substitution;
  therefore, \elpiIn{if (L = L1, S1 = S2)} succeeds and \elpiIn{progress}
  terminates.
\end{proof}

\paragraph{example 1}

The example above shows the new compilation of the unification problem given at
the beginning of \cref{sec:eta}. This time, we see that the the left hand side
$t$ of $\mathbb{P}_1$ has been detected as a \maybeeta and replaced with the
fresh variable $A$. Moreover, \linkStore contains the link $l_1$ connecting $A$
with $\lambda x.X'_{x}$, the compiled version of $t$. After the resolution of
$\mathbb{T}_1$, $A$ is assigned to $f$. Therefore \lhs of $l_1$ is a term with
\elpiIn{con} as constructor. This means that \rhs of $l_1$ is a $\eta$-expansion
and therefore we can unify $\lambda x.f x$ with \rhs, which instantiate $X'_x$
to $f \appsep x$. $l_1$ is removed from \linkStore which is now empty,
\elpiIn{progress} terminate, and the decompilation, will instantiate the $X$
variable to $f$ (which is the $\eta$-contracted version of $\lambda x.f \appsep
x$).

\paragraph{example 2}

@@

\paragraph{example 3}

@@

TODO: we can have $\lambda x. F_x$ in the substitution if we know that $F$ does
not reduce to $T x$ where x is not free in $T$.

>>>>

\noindent
In the base compilation scheme \elpiIn{progress1} is the identity
on both the links and the substitution, so the fixpoint trivially terminates.
Sections \ref{sec:eta} and \ref{sec:beta} add rules to \elpiIn{progress1}
and justify why the don't hinder termination. For brevity we omit the code
that applies the substitution \elpiIn{S1} to all terms in \linkStore.

@@

\begin{lemma}
  \elpiIn{progress} terminates.
\end{lemma}

\begin{proof}[Proof sketch]
  Rules \cref{def:progressetaleft,def:progressetaright} and
  \cref{def:progressetadedup} remove one link from \linkStore, hence they
  cannot be applied indefinitely.
  Moreover each rule only relies on terminating operations such as \Ue,
  $\eta$-contraction, $\eta$-expansion, relocation (a recursive copy of a
  finite term).
  % The addition of rules for \elpiIn{progres1} complicates the function
  % \elpiIn{progress}. We can note, however, that they do not prevent the
  % termination of \elpiIn{progress}. 1) If a link is activated it is removed from
  % \linkStore and the recursive call to \elpiIn{progress} will have a smaller
  % list of links to recurse on. Moreover, link activation only runs terminating
  % instructions (such as unification). 2) If a link is deduplicated, the
  % termination of \elpiIn{progress} is still guaranteed since again we reduce
  % \linkStore and the instructions run by link deduplications are all
  % terminating. 3) If a link is neither activated nor deduplicated, i.e. it
  % remains suspended, then \linkStore remains unchanged like the substitution;
  % therefore, \elpiIn{if (L = L1, S1 = S2)} succeeds and \elpiIn{progress}
  % terminates.
\end{proof}

\paragraph{Example of \progressetaleft}

The example at the beginning of \cref{sec:eta}, once
$\sigma = \{~ A \mapsto f ~\}$, triggers this rule since the link
becomes \linketaM{}{f}{\lambda x.B_{x}} and the \lhs is a constant.
In turn the rule runs $\lambda x.f \appsep x \Ue{} \lambda x.B_{x}$,
resulting in $\sigma = \{~ A \mapsto f ~;~ B_{x} \mapsto f ~\}$.
Decompilation the generates $\rho = \{~ X \mapsto f ~\}$, since
$X$ is mapped to $B$ and
$f$ is the $\eta$-contracted version of $\lambda x.f \appsep x$.

\paragraph{Example of \progressetaright}\marginpar{TODO}

@@

\paragraph{Example of \progressetadedup}

@@

\marginpar{I don't get this TODO}
TODO: we can have $\lambda x. F_x$ in the substitution if we know that $F$ does
not reduce to $T x$ where x is not free in $T$.
